---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# if (Sys.info()["user"] != "bart") {
if (Sys.getenv("MBPWD") != "") {
  options(keyring_backend = "env")
  move2::movebank_store_credentials("move2_user", Sys.getenv("MBPWD"))
} else {
  knitr::opts_chunk$set(eval = FALSE)
}
# }
Sys.setlocale("LC_TIME", "en_US.utf8")
```

```{r setup}
library(moveAcc)
```
The goal of the package is to create a vector representation of acceleration bursts in R. Acceleration bursts are a frequently collected by animal tracking devices and consist of measurements of the acceleration in one till three axis. These measurements are recorded with a fixed frequency for a few seconds. These burst are frequently considered one "instantaneous" sample of the activity or behavior of an animal.

By creating a vector representation it becomes more easy to manipulate these bursts and for example add them to a data frame. 

# Creating a burst

To create a acceleration vector a list of acceleration measurements are combined with the measurement frequency. Each element of the list is a matrix with the columns being the axis the acceleration is measured in and the rows being the repeated measurements.

```{r create}
x <- acc(
  bursts = list(cbind(
    x = sin(1:30 / 10),
    y = cos(1:30 / 10),
    z = 1:30 / 10
  )),
  frequency = units::as_units(40, "Hz")
)
x
```
When printed it reports the average acceleration for each axis of each burst and the measurement frequency. Multiple bursts can be combined with diverse properties can be combined.

```{r combine}
y <- acc(
  bursts = list(
    cbind(x = sin(1:20 / 10), y = cos(1:20 / 10)),
    cbind(x = sin(1:20 / 10 + 2), y = cos(1:20 / 10 + 3))
  ),
  frequency = units::as_units(30, "Hz")
)
y
c(x, y)
```



# Loading data

Naturally most of the time data won't be generated directly in R but rather loaded from local files or databases like movebank. For data from movebank some quick conversion options have been build. These take `move2` objects as an input. For now conversion functions for data from eobs and ornitella tags exists.

Here is an example for eobs tags. When the `as_acc` function is applied to a `move2` it attempts to detect the columns representing the acceleration data. 

```{r eobs}
require(dplyr, quietly = TRUE)
require(move2, quietly = TRUE)
kinkajous_data <- movebank_download_study("Kinkajous on Pipeline Road Panama",
  sensor_type_id = c("acceleration", "gps"), "license-md5" = "6e1743b4b2a919df0f3b4167c94786d9"
)
kinkajous_data <- kinkajous_data %>%
  mutate(acceleration = as_acc(.)) %>%
  select(acceleration)
tail(kinkajous_data)
```
The same can be done with ornitella data.

```{r ornitella}
d <- as.Date("2021-3-3")

lbbg_data <- move2::movebank_download_study("LBBG_ZEEBRUGGE - Lesser black-backed gulls ",
  sensor_type_id = c("gps", "acceleration"),
  timestamp_start = as.POSIXct(d),
  timestamp_end = as.POSIXct(d + 1)
)
lbbg_data <- lbbg_data %>%
  mutate(acceleration = as_acc(.))  %>%
  filter(!is.na(acceleration) | sensor_type_id==653)%>%
  select(acceleration)
lbbg_data
```
```{r stack}
mt_stack(lbbg_data, kinkajous_data)
```
# Exploring data

# Plotting

```{r plot}
plot_time(lbbg_data$acceleration, mt_time(lbbg_data))
```

